Quantifying the importance of an inference model in Bayesian phylogenetics

### 2020-03-30 pirouette_article.tex
Renamed files for thesis
Changes @Giappo

We use the DD tree as shown in Figure \ref{fig:dd_tree},
which has an arbitrary crown age of ten time units
and six taxa.

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/true_tree.png}
  \caption{
    The tree derived from a DD process.
  }
  \label{fig:dd_tree}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{mee}
\bibliography{article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

The code to reproduce our results can be found at  
\url{https://github.com/richelbilderbeek/pirouette_example_30}
and a simplified version is shown here for convenience:

\begin{lstlisting}[language=R]
library(pirouette)

# Create phylogeny
set.seed(314)
phylogeny <- create_exemplary_dd_tree(
  n_taxa = 6, 
  crown_age = 10
)

# Setup pirouette
pir_params <- create_std_pir_params(
  folder_name = "example_30"
)

# Do the runs
pir_out <- pir_run(
  phylogeny = phylogeny,
  pir_params = pir_params
)

# Save
library(ggplot2)
pir_outs <- list()
pir_outs[[1]] <- pir_out
pir_plots(pir_outs) +
  ggtitle("pirouette example 30") +
  ggsave("errors.png", width = 7, height = 7)

pir_save(
  phylogeny = phylogeny,
  pir_params = pir_params,
  pir_out = pir_out,
  folder_name = "example_30"
)
\end{lstlisting}

The DD tree that is generated by this code is as shown in Figure \ref{fig:dd_tree},
which has an arbitrarily chosen crown age of ten time units and six taxa.

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_30/example_30_314/true_tree.png}
  \caption{
    The example tree resulting from a diversity-dependent (DD) simulation.
  }
  \label{fig:dd_tree}
\end{figure}

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% Bibliography
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
% MEE style
\bibliographystyle{pirouette_mee}
\bibliography{pirouette_article}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

### 2020-03-30 pirouette_supplement.tex
Renamed files [WIP]
Changes @Giappo

Until it is on CRAN, and for the most up-to-date version, 
one can download and install the package from \verb;pirouette;'s GitHub 
repository:

\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single
]
remotes::install_github(
  "richelbilderbeek/mcbette"
)
remotes::install_github(
  "richelbilderbeek/pirouette"
)
\end{lstlisting}
To start using \verb;pirouette;, 
load its functions in the global namespace first:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
library(pirouette)
\end{lstlisting}
Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
This can be done from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
beastier::install_beast2()
\end{lstlisting}
For the option to select the best candidate model,
\verb;pirouette; needs the "NS" BEAST2 package [\cite{russel2019model}].
It can be installed from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
mauricer::install_beast2_pkg("NS")
\end{lstlisting}

@@

The code to reproduce this can be found at  
\url{https://github.com/richelbilderbeek/pirouette_example_30}
and a simplified version is shown here for convenience:

\begin{lstlisting}[language=R]
library(pirouette)

# Create phylogeny
set.seed(314)
phylogeny <- create_exemplary_dd_tree(
  n_taxa = 6, 
  crown_age = 10
)

# Setup pirouette
pir_params <- create_std_pir_params(
  folder_name = "example_30"
)

# Do the runs
pir_out <- pir_run(
  phylogeny = phylogeny,
  pir_params = pir_params
)

# Save
library(ggplot2)
pir_outs <- list()
pir_outs[[1]] <- pir_out
pir_plots(pir_outs) +
  ggtitle("pirouette example 30") +
  ggsave("errors.png", width = 7, height = 7)

pir_save(
  phylogeny = phylogeny,
  pir_params = pir_params,
  pir_out = pir_out,
  folder_name = "example_30"
)
\end{lstlisting}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
  \caption{Error distributions of trees, similar to Fig. \ref{fig:example_30} for the main example, but now for a collection of 20 replicate trees. For each setting (true generative, true best candidate, twin generative and twin best candidate), the resulting errors from each replicate pipeline have been merged into a single distribution.}
  \label{fig:replicate_trees}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
  %\caption{Error distributions of trees for 6 taxa, using 20 replicates. This took 16 hours to compute.}
  \caption{Aggregate error distributions for 20 replicates. Here each each true tree has 6 taxa. This took 16 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
  %\caption{Error distributions of trees for 12 taxa, using 10 replicates. This took 19 hours to compute.}
  \caption{Aggregate error distributions for 10 replicates. Here each each true tree has 12 taxa. This took 19 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
  %\caption{Error distributions of trees for 24 taxa, using 20 replicates. This took 28 hours to compute.}
  \caption{Aggregate error distributions for 20 replicates. Here each each true tree has 24 taxa. This took 28 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
  %\caption{Error distributions of trees for 32 taxa, using 5 replicates. This took 7 hours to compute.}
  \caption{Aggregate error distributions for 5 replicates. Here each each true tree has 32 taxa. This took 7 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
  %\caption{Error distributions of trees for 40 taxa, using 2 replicates. This took 8 hours to compute.}
  \caption{Aggregate error distributions for 2 replicates. Here each each true tree has 40 taxa. This took 8 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
  %\caption{Error distributions of trees for a sequence length of 500 nucleotides, using 20 replicates. This took 12 hours to compute.}
  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
  %\caption{Error distributions of trees for a sequence length of 1000 nucleotides (as used in the main example), using 20 replicates. This took 16 hours.}
  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
  %\caption{Error distributions of trees for a sequence length of 2000 nucleotides (as used in the main example), using 10 replicates. This took 11 hours.}
  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
  %\caption{Error distributions of trees when using a Yule prior on a Yule tree, for 10 replicates. This took 9 hours to compute.}
  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
  %\caption{Error distributions of trees when using a Yule prior on a BD tree, for 20 replicates. This took 9 hours to compute.}
  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_23/errors_low.png}
  %\caption{Error distributions of trees using a true tree a DD tree with low likelihood.}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with low likelihood.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_23/errors_mid.png}
  %\caption{Error distributions of trees using a true tree a DD tree with median likelihood.}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_23/errors_high.png}
  %\caption{Error distributions of trees using a true tree a DD tree with high likelihood.}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with high likelihood.}
\end{figure}

@@

%\begin{figure}[H]
%  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
%  \caption{Equal mutation rate and number of mutations, 16 hours (main example)}
%\end{figure}

@@

%begin{figure}[H]
% \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
% \caption{Error distribution for the example tree of the main text which has a per-nucleotide mutation rate of 1.0 / crown age. This took 16 hours to compute.}
%end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age.This took 17 hours to compute.}
\end{figure}

>>>>

Until it is on CRAN, and for the most up-to-date version, 
one can download and install the package from \verb;pirouette;'s GitHub 
repository. We first need the \verb;mcbette; package:
\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single
]
remotes::install_github(
  "richelbilderbeek/mcbette"
)
\end{lstlisting}

Now we can install \verb;pirouette;:
\begin{lstlisting}[
    language = R,
    floatplacement = ht,
    frame = single
]
remotes::install_github(
  "richelbilderbeek/pirouette"
)
\end{lstlisting}
which also installs its dependencies from CRAN.

To start using \verb;pirouette;, 
load its functions in the global namespace first:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
library(pirouette)
\end{lstlisting}
Because \verb;pirouette; calls BEAST2, BEAST2 must be installed. 
This can be done from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
beastier::install_beast2()
\end{lstlisting}
For the option to select the best candidate model,
\verb;pirouette; needs the "NS" BEAST2 package [\cite{russel2019model}].
It can be installed from within R, using:
\begin{lstlisting}[language=R, floatplacement=ht, frame=single]
mauricer::install_beast2_pkg("NS")
\end{lstlisting}

@@

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
  \caption{Aggregate error distributions, similar to Fig. \ref{fig:example_30} for the main example, but now for a collection of 20 replicate trees. For each setting (true generative, true best candidate, twin generative and twin best candidate), the resulting errors from each replicate pipeline have been merged into a single distribution.}
  \label{fig:replicate_trees}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each true tree has 6 taxa. This took 16 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_32/errors.png}
  \caption{Aggregate error distributions for 10 replicates. Here each true tree has 12 taxa. This took 19 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_33/errors.png}
    \caption{Aggregate error distributions for 20 replicates. Here each true tree has 24 taxa. This took 28 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_41/errors.png}
  \caption{Aggregate error distributions for 5 replicates. Here each true tree has 32 taxa. This took 7 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_42/errors.png}
  \caption{Aggregate error distributions for 2 replicates. Here each true tree has 40 taxa. This took 8 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_19/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 500 nucleotides. This took 12 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_28/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each each alignment has a sequence length of 1000 nucleotides (as used in the main example). This took 16 hours to compute.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_34/errors.png}
  \caption{Aggregate error distributions for 10 replicates. Here each each alignment has a sequence length of 2000 nucleotides. This took 11 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_22/errors.png}
  \caption{Aggregate error distributions for 10 replicates. Here each true tree is generated by a Yule process. For the inference we used a Yule tree prior. This took 9 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_26/errors.png}
  \caption{Aggregate error distributions for 20 replicates. Here each true tree is generated by a BD process. For the inference we used instead a Yule tree prior. This took 9 hours to compute.}
\end{figure}

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_23/errors_low.png}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with low likelihood.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_23/errors_mid.png}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with median likelihood.}
\end{figure}

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_23/errors_high.png}
  \caption{Aggregate error distributions for a distribution of trees, where the true trees are DD with high likelihood.}
\end{figure}

@@

@@

\begin{figure}[H]
  \includegraphics[width=\textwidth]{pirouette_example_38/errors.png}
  \caption{Aggregate error distributions for the tree distribution presented in \ref{subsec:distribution} but with a per-nucleotide mutation rate of 1.25 / crown age. This took 17 hours to compute.}
\end{figure}
